\vspace{-0ex}Bench: Evaluating Arbitrary Workloads

### 2017-12-10 paper/submission.tex
Finished adding citations.

\section{Introduction}
Most computer algorithms utilize the storage disk inside of a system. While some access the disk directly, as in the case of data structures like B-Trees and LSM-Trees \cite{}, others may end up accessing them indirectly, perhaps through the means of a paging algorithm in a virtual memory system. These algorithms effectively define a workload which is to be executed by the system. More often than not, the storage medium of a system is the main bottleneck which limits the performance of this workload on any given system \cite{}. This is not hard to believe, particularly because the performance differences between commodity storage media like HDD and SSD pale in comparison to CPU, RAM, and even networking nowadays \cite{}.

@@

Keeping this in mind, it is important to understand that different storage media performs differently for particular types of workloads. Conventional rotating media like hard disk drives perform optimally in a sequential workload \cite{}, whereas flash based storage such as used in solid state drives can perform admirably even in a random-heavy workload \cite{}. However, flash based storage costs more than rotating media storage per giga byte. This means that an engineer designing a system needs to make careful decisions on the inclusion of hard drives, solid state drives and even a combination of both. Driving these decisions is usually a benchmark, highlighting the performance and cost characteristics of the different types of storage media. However, performing such an exhaustive benchmark is a non-trivial task. Different classes of applications produce completely different workloads. For example, a web page may produce small-random oriented workloads whereas a paging algorithm for an operating system might produce a long-sequential one \cite{}.

@@

Most of the information on HDDs presented here is a paraphrasing of \cite{}. For readers looking for a more in-depth and technical explanation for the working of disk drives, chapter 37 in the book provides a solid ground.

@@

Flash memory is a subset of EEPROM (electrically erasable programmable read-only memory). While there are two types of flash memory, namely NAND-flash and NOR-flash, this paper focuses on NAND-flash as it is the more commonly deployed variant \cite{}. Hence, further mentions of flash memory within this paper refer to the use of NAND-flash class of technology.

@@

The main idea behind flash memory is the use of floating-gate transistors \cite{}. Without diving into too much detail, the basic idea involves detecting the presence of a charge in a \textbf{flash cell}. When a certain voltage is applied to parts of the cell, current is allowed to flow through the transistor. We can detect the charge present in the cell by monitoring this current, allowing us to judge the current state of the cell \cite{}. When an even greater voltage is applied, a process known as quantum tunneling allows us to trap an electron in the transistor, allowing us to effectively change the charge of the cell and write a single bit of information to it.

@@

A proposed solution to this problem is the use of management firmware known as a \textbf{flash translation layer (FTL)} \cite{}. An FTL essentially provides a mapping from logical addresses (which are visible to a host system) to physical addresses (which are not visible to a host system) on the flash chip. In an example FTL, if a host were to randomly update a single page in flash memory, the FTL would copy the entire block which contains the requested page over to another place in flash memory, making sure to replace the original page with the modified one. It would then change its mapping table, establishing the new mapping of logical to physical addresses. By doing this, the FTL can avoid erasing blocks frequently, increasing the lifetime of the flash memory. Unfortunately however, in this case, the flash memory incurred a lot more writes than intended, increasing the write latency and decreasing overall throughput. This phenomenon is known as \textbf{write amplification} (WA) and a lot of research has gone into building FTLs which can provide lower write amplification. Regardless of any progress on that front however, write amplification continues to be a major bottleneck for flash memory and prevents random writes (which incur heavy write amplification as explained above) to perform efficiently.

Owing to such characteristics, flash aware software systems have begun using different design paradigms to optimize the use of flash memory. One such example is the use of log based structures wherein all information and operations are performed as appends to a file, making sure that previously written pages are not updated further \cite{}.

An important thing to note is that flash memory performs admirably for random reads when compared to disks because it incurs no overhead from rotations and seeking. However, there are physical challenges which prevent flash based storage to provide the same capacity as disk \cite{}. As a result of that, flash storage is more expensive than disk based storage and requires novel techniques to make it viable for the enterprise \cite{}.

@@

A \textbf{Poisson point process} is a process in which events occur continously and independently at a constant average rate \cite{}. This means that as time $t \Rightarrow \infty$, the average of the intervals between events in a point process converge. Hence, for a system with a known average, \textit{such as a workload}, the poisson point process can model the interval times between a pair of events. An \textbf{exponential distribution} is the probability distribution which describes an interval between a pair of events in a poisson point process \cite{}. Hence, using an exponential distribution, we can accurately model the arrival of work for any given workload.

@@

The reason this system is as such is because industrial applications operate at a huge scale and service thousands of requests at a given time. It is simply not possible for a single machine to keep up with the incoming requests from impatient clients. For a system such as this, the main bottleneck used to be the latency of network connections which connected the frontend to the backend. However, with advances in network technology and the introduction of standards such as \textbf{Infiniband} \cite{}, this is not true anymore. Instead, the main bottleneck for these systems is the storage media, which operate at a fraction of the speed of CPU and RAM.

@@

\subsection{Multithreading}
We implemented the simulation of a distributed system using threads in \textit{bench}. The idea is that we have a \textbf{producer} thread and a \textbf{consumer} thread (which simulate a frontend and a backend respectively) which are both managed by a \textbf{timer} thread. The timer thread works in tandem with the consumer thread to enforce an order of termination between the consumer and the producer thread. If this order is not followed, the application may reach a state where the consumer sleeps for an indefinite amount of amount. This is explained further in the following paragraphs.

@@

>>>>

\section{Introduction}
Most computer algorithms utilize the storage disk inside of a system. While some access the disk directly, as in the case of data structures like B-Trees and LSM-Trees \cite{o1996log}, others may end up accessing them indirectly, perhaps through the means of a paging algorithm in a virtual memory system. These algorithms effectively define a workload which is to be executed by the system. More often than not, the storage medium of a system is the main bottleneck which limits the performance of this workload on any given system \cite{smith1985disk}. This is not hard to believe, particularly because the performance differences between commodity storage media like HDD and SSD pale in comparison to CPU, RAM, and even networking nowadays \cite{boden1995myrinet}.

@@

Keeping this in mind, it is important to understand that different storage media performs differently for particular types of workloads. Conventional rotating media like hard disk drives perform optimally in a sequential workload, whereas flash based storage such as used in solid state drives can perform admirably even in a random-heavy workload. However, flash based storage costs more than rotating media storage per giga byte. This means that an engineer designing a system needs to make careful decisions on the inclusion of hard drives, solid state drives and even a combination of both. Driving these decisions is usually a benchmark, highlighting the performance and cost characteristics of the different types of storage media. However, performing such an exhaustive benchmark is a non-trivial task. Different classes of applications produce completely different workloads. For example, a web page may produce small-random oriented workloads whereas a paging algorithm for an operating system might produce a long-sequential one.

@@

Most of the information on HDDs presented here is a paraphrasing of \cite{osbook}. For readers looking for a more in-depth and technical explanation for the working of disk drives, chapter 37 in the book provides a solid ground.

@@

Flash memory is a subset of EEPROM (electrically erasable programmable read-only memory). While there are two types of flash memory, namely NAND-flash and NOR-flash, this paper focuses on NAND-flash as it is the more commonly deployed variant \cite{micheloni2010nand}. Hence, further mentions of flash memory within this paper refer to the use of NAND-flash class of technology.

@@

The main idea behind flash memory is the use of floating-gate transistors \cite{micheloni2010nand}. Without diving into too much detail, the basic idea involves detecting the presence of a charge in a \textbf{flash cell}. When a certain voltage is applied to parts of the cell, current is allowed to flow through the transistor. We can detect the charge present in the cell by monitoring this current, allowing us to judge the current state of the cell \cite{micheloni2010nand}. When an even greater voltage is applied, a process known as quantum tunneling allows us to trap an electron in the transistor, allowing us to effectively change the charge of the cell and write a single bit of information to it.

@@

A proposed solution to this problem is the use of management firmware known as a \textbf{flash translation layer (FTL)} \cite{kim2002space}. An FTL essentially provides a mapping from logical addresses (which are visible to a host system) to physical addresses (which are not visible to a host system) on the flash chip. In an example FTL, if a host were to randomly update a single page in flash memory, the FTL would copy the entire block which contains the requested page over to another place in flash memory, making sure to replace the original page with the modified one. It would then change its mapping table, establishing the new mapping of logical to physical addresses. By doing this, the FTL can avoid erasing blocks frequently, increasing the lifetime of the flash memory. Unfortunately however, in this case, the flash memory incurred a lot more writes than intended, increasing the write latency and decreasing overall throughput. This phenomenon is known as \textbf{write amplification} (WA) and a lot of research has gone into building FTLs which can provide lower write amplification. Regardless of any progress on that front however, write amplification continues to be a major bottleneck for flash memory and prevents random writes (which incur heavy write amplification as explained above) to perform efficiently.

Owing to such characteristics, flash aware software systems have begun using different design paradigms to optimize the use of flash memory. One such example is the use of log based structures wherein all information and operations are performed as appends to a file, making sure that previously written pages are not updated further \cite{rosenblum1992design}.

An important thing to note is that flash memory performs admirably for random reads when compared to disks because it incurs no overhead from rotations and seeking. However, there are physical challenges which prevent flash based storage to provide the same capacity as disk \cite{micheloni2010nand}. As a result of that, flash storage is more expensive than disk based storage and requires novel techniques to make it viable for the enterprise \cite{colgrove2015purity}.

@@

A \textbf{Poisson point process} is a process in which events occur continously and independently at a constant average rate \cite{kingman1993poisson}. This means that as time $t \Rightarrow \infty$, the average of the intervals between events in a point process converge. Hence, for a system with a known average, \textit{such as a workload}, the poisson point process can model the interval times between a pair of events. An \textbf{exponential distribution} is the probability distribution which describes an interval between a pair of events in a poisson point process \cite{buzen1973computational}. Hence, using an exponential distribution, we can accurately model the arrival of work for any given workload.

@@

The reason this system is as such is because industrial applications operate at a huge scale and service thousands of requests at a given time. It is simply not possible for a single machine to keep up with the incoming requests from impatient clients. For a system such as this, the main bottleneck used to be the latency of network connections which connected the frontend to the backend. However, with advances in network technology and the introduction of standards such as \textbf{Infiniband} \cite{pfister2001introduction}, this is not true anymore. Instead, the main bottleneck for these systems is the storage media, which operate at a fraction of the speed of CPU and RAM.

@@

\subsection{Multithreading}
We implemented the simulation of a distributed system using the \texttt{pthreads} library in \textit{bench}. The idea is that we have a \textbf{producer} thread and a \textbf{consumer} thread (which simulate a frontend and a backend respectively) which are both managed by a \textbf{timer} thread. The timer thread works in tandem with the consumer thread to enforce an order of termination between the consumer and the producer thread. If this order is not followed, the application may reach a state where the consumer sleeps for an indefinite amount of time. This is explained further in the following paragraphs.

@@

\section{Evaluations}

\section{Conclusion}

\bibliographystyle{plain}
\bibliography{cites}
